---
title: "dataframes"
author: "C. Durso"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Read in data

General text files may be read in using *read.table* using appropriate arguments. CSV format, an "industry standard" typically uses *read.csv*

```{r}
dat<-read.csv("worms.csv")
# Get just a few rows:
dat.b<-read.csv("worms.csv",nrows=5)
```

## Inspecting the data

You can view dataframes by double clicking in RStudio. There are some standard summary tools.

```{r}
dat
dim(dat)
names(dat)
str(dat)
summary(dat)
head(dat)
dat[2,4]
```

## Select columns

Leaving the row index blank and providing column indices allows selection of a subset of the columns.

```{r}
dat[,3]
dat$Vegetation
dat[,2:4]

```

## Select rows

Leaving the column index blank and providing a vector of indices or boolean values allows selection of a subset of the rows.

```{r}
dat[1:4,]
damp<-dat[dat$Damp,]
summary(damp)
flat<-dat[dat$Slope<4.5,]
summary(flat)
```

## Sorting

The function *sort* applied to a vector returns the sorted vector.

```{r}
sort(dat[,2])
dat[,2]
```

To sort a dataframe on a column, apply *order* to the column. It returns a vector of the index of each element of the column after a stable sort.

```{r}
ind<-order(dat$Area)
dat$Area[ind]
dat.ordered<-dat[ind,] # reorder rows
dat.ordered$Area
head(dat.ordered)
ind<-order(dat.ordered$Slope)
dat.ordered<-dat.ordered[ind,] # ordered by slope, and by area within slope
head(dat.ordered)
```

Alternatively, the *arrange* function may be used.

```{r}
dat.ordered2<-arrange(dat,Slope,Area)
head(dat.ordered2)
```

## Saving

If you have built data that you would like to use elsewhere, you can save it:

```{r}
save(dat,file="worm_data.RData")
rm(dat)
ls() 
load("worm_data.RData")
ls()
str(dat) # it's back
```

You can save plots by plotting from the Console then exporting.

```{r}
qplot(dat$Soil.pH,dat$Worm.density,main="Worm Density by pH", xlab = "pH",
     ylab = "worm density")
```

or through code.

```{r}
png(file = "myplot.png", bg = "transparent")
qplot(dat$Soil.pH,dat$Worm.density,main="Worm Density by pH", xlab = "pH",
     ylab = "worm density")
dev.off()

```
Check in Files tab. 